Preparación de datos

### 2024-12-16 04-analisis.qmd
add covid vars
significant effect with:
- it is more important to give priority to economic activity than to the health of the population
- receive State benefits (Food Box, Emergency Family Income, Contribution for the Middle Class or other)

# generate analytical sample
df_study1 <- 
  df_study_long_t7_comunas %>%
  select(idencuesta,ola,comuna, ponderador_long_total,just_educ, des_perc, merit_effort, merit_talent, education=educ,quintil1:sexo) %>% 
  na.omit() %>% 
  mutate(ola = as.factor(ola),
         ola_num = as.numeric(ola),
         ola_2=as.numeric(ola)^2,
         sexo=as.factor(sexo)
         )
# Hipotesis
h0 <- "ola"
h1 <- "ola_num+ola_2"
h2 <- "ola_num+ola_2+education+quintil1+ess+edad+sexo"
h3 <- "ola_num+ola_2+education+quintil1+ess+edad+sexo+pos_id"
h4 <- "ola_num+ola_2+education+quintil1+ess+edad+sexo+pos_id+des_perc"
h5 <- "ola_num+ola_2+education+quintil1+ess+edad+sexo+pos_id+des_perc+merit_talent"
h6 <- "ola_num+ola_2+education+quintil1+ess+edad+sexo+pos_id+des_perc+merit_talent+merit_effort"

# A. Education distributive justice
educ.null <- lmer(formula(paste0("just_educ~","1 + (1|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ0 <- lmer(formula(paste0("just_educ~",h0,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ1 <- lmer(formula(paste0("just_educ~",h1,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ2 <- lmer(formula(paste0("just_educ~",h2,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ3 <- lmer(formula(paste0("just_educ~",h3,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ4 <- lmer(formula(paste0("just_educ~",h4,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ5 <- lmer(formula(paste0("just_educ~",h5,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ6 <- lmer(formula(paste0("just_educ~",h6,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)

@@

omit <- "(education)|(ess)"
texreg::knitreg(list(educ0, educ1, educ3, educ4, educ5, educ6),
                custom.note = "*** p < 0.001; ** p < 0.01; * p < 0.05.
Note: Model 3 to 6 are controlled by educational level, and subjective social status with no significant effects.", 
                omit.coef = omit,
                custom.coef.names = c("Intercept",
                                      "Wave 2017",
                                      "Wave 2018",
                                      "Wave 2019",
                                      "Wave 2022",
                                      "Wave 2023",
                                      "Wave",
                                      "Wave^2",
#                                      "High School <br> <i>(Ref. Primary)</i>",
#                                      "Technical",
#                                      "Universitary",
                                      "Quintile 2",
                                      "Quintile 3",
                                      "Quintile 4",
                                      "Quintile 5",
                                      "Quintile (No information)",
#                                      "Subjective Social Status",
                                      "Age 30-49",
                                      "Age 50-64",
                                      "Age 65 or more",
                                      "Gender (Ref. Male)",
                                      "Center",
                                      "Right",
                                      "Does not identify",
                                      "Inequality perception",
                                      "Merit: Talent",
                                      "Merit: Effort" ),
                groups = list("Wave (Ref.= Wave 2016)" = 2:6,
                  "Household Income (Ref.= Quintile 1)" = 9:13,
                  "Age (Ref.= 18-29)" = 14:16,
                  "Pol. pos (Ref.= Left)" = 18:20
                  ),
                include.loglik = FALSE,
                include.aic = FALSE,
                custom.gof.names = c("BIC",
                                     "Num. obs.",
                                     "Num. groups: Individuals",
                                     "Var: Individuals (Intercept)",
                                     "Var: Residual"),
               center = TRUE,
               caption.above = TRUE,
               scalebox = 0.765,
               float.pos = "!ht", 
               booktabs = FALSE, 
               use.packages = FALSE, 
               include.rs = FALSE, 
               include.rmse = FALSE,
               include.ci = FALSE)

@@

df_study1 <- 
  df_study_long_t7_comunas %>%
  select(idencuesta,ola,comuna, ponderador_long_total,just_educ, des_perc, merit_effort, merit_talent,education=educ,quintil1:sexo) %>% 
  na.omit() %>% 
  mutate(ola = as.numeric(ola),
         ola_2=as.numeric(ola)^2,
         sexo=as.factor(sexo)
 #        edad=as.numeric(edad)
         #pos_id=as.numeric(pos_id)
         )

# Hipotesis

h6 <- "ola+ola_2+education+quintil1+ess+edad+sexo+pos_id+des_perc+merit_talent+merit_effort"
h1b <- "ola+ola_2+education+quintil1+ess+edad+sexo+pos_id+des_perc+merit_talent+merit_effort+merit_effort*ola"
h3b <- "ola+ola_2+education+quintil1+ess+edad+sexo+pos_id+des_perc+merit_talent+merit_effort+merit_talent*ola"

@@

# table 3: interacciones
omit <- "(education)|(ess)|(quintil1)|(edad)|(sexo)|(pos_id)|(des_perc)"
texreg::knitreg(list(educ6,educ3,educ1),
                custom.note = "\\item *** p < 0.001; ** p < 0.01; * p < 0.05. Note: All the models are controlled by educational level, income quintile, subjective social status, age, gender, political position, and inequality perception",
                omit.coef = omit,
                include.loglik = FALSE,
                include.aic = FALSE,
                custom.coef.names = c("Intercept",
                                      "Wave",
                                      "Wave^2",
                                      "Merit: Talent",
                                      "Merit: Effort",
                                      "Wave * Merit: Talent",
                                      "Wave * Merit: Effort"
                                      ),
                custom.gof.names = c("BIC",
                                     "Num. obs.",
                                     "Num. groups: Individuals",
                                     "Var: Individuals (Intercept)",
                                     "Var: Residual",
                                     "Var: Individuals Wave",
                                     "Var: Individuals Merit: Talent",
                                     "Cov: Individuals (Intercept) Wave",
                                     "Cov: Individuals (Intercept) Merit: Talent",
                                     "Cov: Individuals Wave Merit: Talent",
                                     "Var: Individuals Merit: Effort",
                                     "Cov: Individuals (Intercept) Merit: Effort",
                                     "Cov: Individuals Wave Merit: Effort"),  
               center = TRUE,
               caption.above = TRUE,
               scalebox = 0.765,
               float.pos = "!ht", 
               booktabs = TRUE, 
               use.packages = FALSE, 
               include.rs = FALSE, 
               include.rmse = FALSE,
               include.ci = FALSE,
               threeparttable = TRUE)
```

>>>>

# generate analytical sample
df_study1 <- 
  df_study_long_t7_comunas %>%
  select(idencuesta,ola,comuna, ponderador_long_total,just_educ, des_perc, merit_effort, merit_talent, education=educ,quintil1:sexo, covid_economia, covid_beneficios) %>% 
  na.omit() %>% 
  mutate(ola = as.factor(ola),
         ola_num = as.numeric(ola),
         ola_2=as.numeric(ola)^2,
         sexo=as.factor(sexo)
         )

df_study1$covid_beneficios<- car::recode(df_study1$covid_beneficios, c("1=2; 2=1"))

# Hipotesis
h0 <- "ola"
h1 <- "ola_num+ola_2"
h2 <- "ola_num+ola_2+education+quintil1+ess+edad+sexo"
h3 <- "ola_num+ola_2+education+quintil1+ess+edad+sexo+pos_id"
h4 <- "ola_num+ola_2+education+quintil1+ess+edad+sexo+pos_id+des_perc"
h5 <- "ola_num+ola_2+education+quintil1+ess+edad+sexo+pos_id+des_perc+merit_talent"
h6 <- "ola_num+ola_2+education+quintil1+ess+edad+sexo+pos_id+des_perc+merit_talent+merit_effort"
h7 <- "ola_num+ola_2+education+quintil1+ess+edad+sexo+pos_id+des_perc+merit_talent+merit_effort+covid_economia"
h8 <- "ola_num+ola_2+education+quintil1+ess+edad+sexo+pos_id+des_perc+merit_talent+merit_effort+covid_economia+covid_beneficios"

# A. Education distributive justice
educ.null <- lmer(formula(paste0("just_educ~","1 + (1|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ0 <- lmer(formula(paste0("just_educ~",h0,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ1 <- lmer(formula(paste0("just_educ~",h1,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ2 <- lmer(formula(paste0("just_educ~",h2,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ3 <- lmer(formula(paste0("just_educ~",h3,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ4 <- lmer(formula(paste0("just_educ~",h4,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ5 <- lmer(formula(paste0("just_educ~",h5,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ6 <- lmer(formula(paste0("just_educ~",h6,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ7 <- lmer(formula(paste0("just_educ~",h7,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)
educ8 <- lmer(formula(paste0("just_educ~",h8,"+(1|idencuesta)")),data = df_study1, weights = ponderador_long_total)

@@

omit <- "(education)|(ess)"
texreg::knitreg(list(educ0, educ1, educ3, educ4, educ5, educ6,educ7,educ8),
                custom.note = "*** p < 0.001; ** p < 0.01; * p < 0.05.
Note: Model 3 to 6 are controlled by educational level, and subjective social status with no significant effects.", 
                omit.coef = omit,
                custom.coef.names = c("Intercept",
                                      "Wave 2017",
                                      "Wave 2018",
                                      "Wave 2019",
                                      "Wave 2022",
                                      "Wave 2023",
                                      "Wave",
                                      "Wave^2",
#                                      "High School <br> <i>(Ref. Primary)</i>",
#                                      "Technical",
#                                      "Universitary",
                                      "Quintile 2",
                                      "Quintile 3",
                                      "Quintile 4",
                                      "Quintile 5",
                                      "Quintile (No information)",
#                                      "Subjective Social Status",
                                      "Age 30-49",
                                      "Age 50-64",
                                      "Age 65 or more",
                                      "Gender (Ref. Male)",
                                      "Center",
                                      "Right",
                                      "Does not identify",
                                      "Inequality perception",
                                      "Merit: Talent",
                                      "Merit: Effort" ,
                                      "Covid: economy more important than health",
                                      "Covid: receive state benefits"),
                groups = list("Wave (Ref.= Wave 2016)" = 2:6,
                  "Household Income (Ref.= Quintile 1)" = 9:13,
                  "Age (Ref.= 18-29)" = 14:16,
                  "Pol. pos (Ref.= Left)" = 18:20
                  ),
                include.loglik = FALSE,
                include.aic = FALSE,
                custom.gof.names = c("BIC",
                                     "Num. obs.",
                                     "Num. groups: Individuals",
                                     "Var: Individuals (Intercept)",
                                     "Var: Residual"),
               center = TRUE,
               caption.above = TRUE,
               scalebox = 0.765,
               float.pos = "!ht", 
               booktabs = FALSE, 
               use.packages = FALSE, 
               include.rs = FALSE, 
               include.rmse = FALSE,
               include.ci = FALSE)

@@

df_study1 <- 
  df_study_long_t7_comunas %>%
  select(idencuesta,ola,comuna, ponderador_long_total,just_educ, des_perc, merit_effort, merit_talent,education=educ,quintil1:sexo, covid_economia, covid_beneficios) %>% 
  na.omit() %>% 
  mutate(ola = as.numeric(ola),
         ola_2=as.numeric(ola)^2,
         sexo=as.factor(sexo)
 #        edad=as.numeric(edad)
         #pos_id=as.numeric(pos_id)
         )

df_study1$covid_beneficios<- car::recode(df_study1$covid_beneficios, c("1=2; 2=1"))
# Hipotesis

h6 <- "ola+ola_2+education+quintil1+ess+edad+sexo+pos_id+des_perc+merit_talent+merit_effort+covid_economia+covid_beneficios"
h1b <- "ola+ola_2+education+quintil1+ess+edad+sexo+pos_id+des_perc+merit_talent+merit_effort+covid_economia+covid_beneficios+merit_effort*ola"
h3b <- "ola+ola_2+education+quintil1+ess+edad+sexo+pos_id+des_perc+merit_talent+merit_effort+covid_economia+covid_beneficios+merit_talent*ola"

@@

# table 3: interacciones
omit <- "(education)|(ess)|(quintil1)|(edad)|(sexo)|(pos_id)|(des_perc)|(covid_beneficios)|(covid_economia)"
texreg::knitreg(list(educ6,educ3,educ1),
                custom.note = "\\item *** p < 0.001; ** p < 0.01; * p < 0.05. Note: All the models are controlled by educational level, income quintile, subjective social status, age, gender, political position, and inequality perception",
                omit.coef = omit,
                include.loglik = FALSE,
                include.aic = FALSE,
                custom.coef.names = c("Intercept",
                                      "Wave",
                                      "Wave^2",
                                      "Merit: Talent",
                                      "Merit: Effort",
                                      "Wave * Merit: Talent",
                                      "Wave * Merit: Effort"
                                      ),
                custom.gof.names = c("BIC",
                                     "Num. obs.",
                                     "Num. groups: Individuals",
                                     "Var: Individuals (Intercept)",
                                     "Var: Residual",
                                     "Var: Individuals Wave",
                                     "Var: Individuals Merit: Talent",
                                     "Cov: Individuals (Intercept) Wave",
                                     "Cov: Individuals (Intercept) Merit: Talent",
                                     "Cov: Individuals Wave Merit: Talent",
                                     "Var: Individuals Merit: Effort",
                                     "Cov: Individuals (Intercept) Merit: Effort",
                                     "Cov: Individuals Wave Merit: Effort"),  
               center = TRUE,
               caption.above = TRUE,
               scalebox = 0.765,
               float.pos = "!ht", 
               booktabs = TRUE, 
               use.packages = FALSE, 
               include.rs = FALSE, 
               include.rmse = FALSE,
               include.ci = FALSE,
               threeparttable = TRUE)
```

### 2024-12-16 production/prod_prep.Rmd
add covid vars
significant effect with:
- it is more important to give priority to economic activity than to the health of the population
- receive State benefits (Food Box, Emergency Family Income, Contribution for the Middle Class or other)

# Attitudes toward inmigrants__________________________________________________
elsoc_long_2016_2023 <- 
elsoc_long_2016_2023 %>% 
  #create new variables
  mutate(just_educ=d02_02,
         just_pension=d02_01,
         just_salud=d02_03,
         des_perc=c18_11,
         sexo=m0_sexo,
         merit_effort=c18_09,
         merit_talent=c18_10,
         ahead_family=r12_02,
         ahead_educ=r12_03,
         ahead_ambition=r12_04,
         ahead_hardwork=r12_05) %>% 
  sjlabelled::drop_labels(., drop.na = FALSE)

@@

sjmisc::frq(elsoc_long_2016_2023$pos_id)

# Select variables______________________________________________________________

df_study1 <- 
elsoc_long_2016_2023 %>% 
  filter(muestra == 1) %>% #keep original sample (vs refresh)
  select(idencuesta,ola,region,region_cod,comuna,comunacod=comuna_cod,
         muestra,
         ponderador_long_total, segmento, estrato,
         sexo,
         # cuestion_mig,
         # starts_with("in_"),
         just_educ,
         just_pension,
         just_salud,
         des_perc,
         merit_effort,
         merit_talent,
         ahead_family,
         ahead_educ,
         ahead_ambition,
         ahead_hardwork,
         educ,educyear,
         # ing_pc,
         # quintil,
         quintil1,
         ess,
         # ess_inm,
         # nation_t1,
         # know_inm,
         # know_inm_bi,
         # frien_inm_bi_t4,
         # frien_inm_bi_t1,
         # know_inm_t1t4,
         # frien_inm,
         # frien_inm_bi,
         # frien_inm_t1t4,
         pos_id,
         edad)

@@

@@

# reshape from long to wide
pacman::p_load(datasets,data.table)
df_study1_long <- data.table::melt.data.table(data.table::setDT(df_study1_wide),
              id.vars = c("idencuesta","muestra"),
              variable.name = c("ola"),
              measure = patterns("^comuna_","^comunacod_","^ponderador_long_total", "^segmento", "^estrato",
                                 "^just_educ_", "^just_pension_", "^just_salud_", "^des_perc_",
                                 "^merit_effort_","^merit_talent_",
                                 "^ahead_family_", "^ahead_educ_", "^ahead_ambition_", "^ahead_hardwork_",
                                 "^educ_","^educyear_","^quintil1_","^ess_",
                                 "^pos_id_",
                                 "^edad_","^sexo_"
                                 ),
              value.name = c("comuna","comunacod","ponderador_long_total","segmento","estrato",
                             "just_educ", "just_pension","just_salud", "des_perc",
                             "merit_effort","merit_talent",
                             "ahead_family","ahead_educ","ahead_ambition","ahead_hardwork",
                             "educ","educyear","quintil1","ess",
                             "pos_id",
                             "edad","sexo"),
              na.rm = F,value.factor = TRUE
              )

>>>>

# Attitudes toward inmigrants__________________________________________________
elsoc_long_2016_2023 <- 
elsoc_long_2016_2023 %>% 
  #create new variables
  mutate(just_educ=d02_02,
         just_pension=d02_01,
         just_salud=d02_03,
         des_perc=c18_11,
         sexo=m0_sexo,
         merit_effort=c18_09,
         merit_talent=c18_10,
         ahead_family=r12_02,
         ahead_educ=r12_03,
         ahead_ambition=r12_04,
         ahead_hardwork=r12_05,
         covid_conocidos=r03_09,
         covid_economia=c37_09,
         covid_cuarentena=c37_10,
         covid_propio=s31,
         covid_alimentos=m62,
         covid_beneficios=m63_01,
         covid_retirouno=m63_02,
         covid_retirodos=m63_03
         ) %>% 
  sjlabelled::drop_labels(., drop.na = FALSE)

@@

sjmisc::frq(elsoc_long_2016_2023$pos_id)
```

```{r}
# Select variables______________________________________________________________

df_study1 <- 
elsoc_long_2016_2023 %>% 
  filter(muestra == 1) %>% #keep original sample (vs refresh)
  select(idencuesta,ola,region,region_cod,comuna,comunacod=comuna_cod,
         muestra,
         ponderador_long_total, segmento, estrato,
         sexo,
         # cuestion_mig,
         # starts_with("in_"),
         just_educ,
         just_pension,
         just_salud,
         des_perc,
         merit_effort,
         merit_talent,
         ahead_family,
         ahead_educ,
         ahead_ambition,
         ahead_hardwork,
         educ,educyear,
         # ing_pc,
         # quintil,
         quintil1,
         ess,
         covid_conocidos,
         covid_economia,
         covid_cuarentena,
         covid_propio,
         covid_alimentos,
         covid_beneficios,
         covid_retirouno,
         covid_retirodos,
         # ess_inm,
         # nation_t1,
         # know_inm,
         # know_inm_bi,
         # frien_inm_bi_t4,
         # frien_inm_bi_t1,
         # know_inm_t1t4,
         # frien_inm,
         # frien_inm_bi,
         # frien_inm_t1t4,
         pos_id,
         edad)

@@

df_study1_wide$covid_conocidos_1 <-df_study1_wide$covid_conocidos_5 #conocidos covid
df_study1_wide$covid_conocidos_2 <-df_study1_wide$covid_conocidos_5
df_study1_wide$covid_conocidos_3 <-df_study1_wide$covid_conocidos_5
df_study1_wide$covid_conocidos_4 <-df_study1_wide$covid_conocidos_5
df_study1_wide$covid_conocidos_6 <-df_study1_wide$covid_conocidos_5
df_study1_wide$covid_conocidos_7 <-df_study1_wide$covid_conocidos_5

df_study1_wide$covid_economia_1 <-df_study1_wide$covid_economia_5 #Importancia economia sobre salud
df_study1_wide$covid_economia_2 <-df_study1_wide$covid_economia_5
df_study1_wide$covid_economia_3 <-df_study1_wide$covid_economia_5
df_study1_wide$covid_economia_4 <-df_study1_wide$covid_economia_5
df_study1_wide$covid_economia_6 <-df_study1_wide$covid_economia_5
df_study1_wide$covid_economia_7 <-df_study1_wide$covid_economia_5

df_study1_wide$covid_cuarentena_1 <-df_study1_wide$covid_cuarentena_5 #importancia cuarentena
df_study1_wide$covid_cuarentena_2 <-df_study1_wide$covid_cuarentena_5
df_study1_wide$covid_cuarentena_3 <-df_study1_wide$covid_cuarentena_5
df_study1_wide$covid_cuarentena_4 <-df_study1_wide$covid_cuarentena_5
df_study1_wide$covid_cuarentena_6 <-df_study1_wide$covid_cuarentena_5
df_study1_wide$covid_cuarentena_7 <-df_study1_wide$covid_cuarentena_5

df_study1_wide$covid_propio_1 <-df_study1_wide$covid_propio_5 #Diagnostico covid propio
df_study1_wide$covid_propio_2 <-df_study1_wide$covid_propio_5
df_study1_wide$covid_propio_3 <-df_study1_wide$covid_propio_5
df_study1_wide$covid_propio_4 <-df_study1_wide$covid_propio_5
df_study1_wide$covid_propio_6 <-df_study1_wide$covid_propio_5
df_study1_wide$covid_propio_7 <-df_study1_wide$covid_propio_5

df_study1_wide$covid_alimentos_1 <-df_study1_wide$covid_alimentos_5 # Dificultades conseguir alimentos
df_study1_wide$covid_alimentos_2 <-df_study1_wide$covid_alimentos_5
df_study1_wide$covid_alimentos_3 <-df_study1_wide$covid_alimentos_5
df_study1_wide$covid_alimentos_4 <-df_study1_wide$covid_alimentos_5
df_study1_wide$covid_alimentos_6 <-df_study1_wide$covid_alimentos_5
df_study1_wide$covid_alimentos_7 <-df_study1_wide$covid_alimentos_5

df_study1_wide$covid_beneficios_1 <-df_study1_wide$covid_beneficios_5 #Recibe beneficios
df_study1_wide$covid_beneficios_2 <-df_study1_wide$covid_beneficios_5
df_study1_wide$covid_beneficios_3 <-df_study1_wide$covid_beneficios_5
df_study1_wide$covid_beneficios_4 <-df_study1_wide$covid_beneficios_5
df_study1_wide$covid_beneficios_6 <-df_study1_wide$covid_beneficios_5
df_study1_wide$covid_beneficios_7 <-df_study1_wide$covid_beneficios_5

df_study1_wide$covid_retirouno_1 <-df_study1_wide$covid_retirouno_5 #Primer retiro afp
df_study1_wide$covid_retirouno_2 <-df_study1_wide$covid_retirouno_5
df_study1_wide$covid_retirouno_3 <-df_study1_wide$covid_retirouno_5
df_study1_wide$covid_retirouno_4 <-df_study1_wide$covid_retirouno_5
df_study1_wide$covid_retirouno_6 <-df_study1_wide$covid_retirouno_5
df_study1_wide$covid_retirouno_7 <-df_study1_wide$covid_retirouno_5

df_study1_wide$covid_retirodos_1 <-df_study1_wide$covid_retirodos_5 #Segundo retiro afp
df_study1_wide$covid_retirodos_2 <-df_study1_wide$covid_retirodos_5
df_study1_wide$covid_retirodos_3 <-df_study1_wide$covid_retirodos_5
df_study1_wide$covid_retirodos_4 <-df_study1_wide$covid_retirodos_5
df_study1_wide$covid_retirodos_6 <-df_study1_wide$covid_retirodos_5
df_study1_wide$covid_retirodos_7 <-df_study1_wide$covid_retirodos_5

@@

# reshape from long to wide
pacman::p_load(datasets,data.table)
df_study1_long <- data.table::melt.data.table(data.table::setDT(df_study1_wide),
              id.vars = c("idencuesta","muestra"),
              variable.name = c("ola"),
              measure = patterns("^comuna_","^comunacod_","^ponderador_long_total", "^segmento", "^estrato",
                                 "^just_educ_", "^just_pension_", "^just_salud_", "^des_perc_",
                                 "^merit_effort_","^merit_talent_",
                                 "^ahead_family_", "^ahead_educ_", "^ahead_ambition_", "^ahead_hardwork_",
                                 "^covid_conocidos_", "^covid_economia_", "^covid_cuarentena_", "^covid_propio_",
                                 "^covid_alimentos_", "^covid_beneficios_", "^covid_retirouno_", "^covid_retirodos_",
                                 "^educ_","^educyear_","^quintil1_","^ess_",
                                 "^pos_id_",
                                 "^edad_","^sexo_"
                                 ),
              value.name = c("comuna","comunacod","ponderador_long_total","segmento","estrato",
                             "just_educ", "just_pension","just_salud", "des_perc",
                             "merit_effort","merit_talent",
                             "ahead_family","ahead_educ","ahead_ambition","ahead_hardwork",
                             "covid_conocidos", "covid_economia", "covid_cuarentena", "covid_propio",
                             "covid_alimentos", "covid_beneficios", "covid_retirouno", "covid_retirodos",
                             "educ","educyear","quintil1","ess",
                             "pos_id",
                             "edad","sexo"),
              na.rm = F,value.factor = TRUE
              )
